---
title: "GEOG5927 Predictive Analytics : Introduction"
author: "Roger Beecham"
date: "last updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, "slides-theme.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      slideNumberFormat: ''
      titleSlideClass: [center, middle]
---

```{r setup, include=FALSE, echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(comment = "")
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)
```

class: center middle

.center[
<img src = "img/geog5927.png", width = 40%></img>
]

---

## Module content and philosophy


* Spatial modelling to simulate and predict consumer behaviour
  + Exploratory analysis
  + Microsimulation
  + Agent-based modelling

<br>

* Research case studies to evaluate modelling techniques in practice
  + Practical sessions
  + Individual data science report
  + Group data science presentation

---

## Outcomes

By the end of this module you should be able to:

1. **Explain** and **critically evaluate** the role of spatial analytics in simulating and predicting consumer behaviours

2. **Apply** geocomputational modelling and simulation techniques on real data sets

3. **Devise** and **employ** spatial modelling tools to address business problems, presenting and justifying recommendations in an appropriate context

---

## Module team


```{r team, echo=FALSE, out.width="100%"}
team <- tibble::tibble(
  img = "",
  name = c("Roger Beecham", "Jiaqi Ge", "Nik Malleson", "Rachel Oldroyd"),
  role = c("Convenor", "Lecturer", "Lecturer", "Lecturer"),
  sessions = c("1,2,4,5", "3,5", "3,5", "4,5"),
  assignments = c("1", "2", "NA", "NA")
)
kbl(team) %>%
  row_spec(0, background = "#ffffff", font_size = 18) %>%
  row_spec(1:4, background = "#ffffff", font_size = 20) %>%
  column_spec(1, image = spec_image(
    c("img/profile_roger.jpg", "img/profile_jg.jpg", "img/profile_nik.jpg", "img/profile_rachel.jpeg"), 250, 250))
```

---
## Scheduling


```{r schedule, echo=FALSE, out.width="100%"}
schedule <- tibble::tibble(
  session = c("1", "2", "3", "4", "5"),
  wc =c("21 Jun", "28 Jun", "5 Jul", "12 Jul", "19 Jul"),
  lecture = c("mon 1400-1500", "mon 1400-1500", "mon 1400-1500", "mon 1400-1500", "mon 1400-1500"),
  lab = c("wed 0900-1200", "wed 0900-1200", "wed 0900-1200", "wed 0900-1200", "NA - pres"),
  recap = c("fri 0900-1000", "fri 0900-1000", "fri 0900-1000", "fri 0900-1000", "NA")
)
kbl(schedule) %>%
  row_spec(0, background = "#ffffff", font_size = 18 ) %>%
  row_spec(1:5, background = "#ffffff", font_size = 20)
```

---

## Outline : Lectures


```{r outline-lectures, echo=FALSE, out.width="100%"}
outline <- tibble::tibble(
  session = c("1", "2", "3", "4", "5"),
  wc =c("21 Jun", "28 Jun", "5 Jul", "12 Jul", "19 Jul"),
  academic=c("RB","RB", "JG/NM", "RO/RB", "RB/JG/NM/RO"),
  lecture = c("Predictive analytics & microsimulation", "Response modelling and targeted marketing", "Behavioural and agent-based models",
  "Guest lecture", "Wrap-up")
)
kbl(outline) %>%
  row_spec(0, background = "#ffffff", font_size = 18 ) %>%
  row_spec(1:5, background = "#ffffff", font_size=20)
```
<br>
.small-font[
RB: Roger Beecham, JG: Jiaqi Ge, NM: Nik Malleson, RO: Rachel Oldroyd
]
---

## Assessment


* Assignment 1: 75%
  + Individual data analysis based on labs 1 and 2
  + 2,000 words, 4 figures
  + Thursday 15th July 2021 by 2pm

<br>


* Assignment 2: 25%
  + Group presentations based on practical 3
  + Presentations held Wednesday 21st July 2021

---

## Assessment


```{r outline-lectures-ass, echo=FALSE, out.width="100%"}
outline <- tibble::tibble(
  session = c("1", "2", "3", "4", "5"),
  wc =c("21 Jun", "28 Jun", "5 Jul", "12 Jul", "19 Jul"),
  academic=c("RB","RB", "JG/NM", "RO/RB", "RB/JG/NM/RO"),
  lecture = c("Predictive analytics & microsimulation", "Response modelling and targeted marketing", "Behavioural and agent-based models",
  "Guest lecture", "Wrap-up"),
  deadline=c("","","","Ass 1","Ass 2")
)
kbl(outline) %>%
  row_spec(0, background = "#ffffff", color="#616161", font_size = 18 ) %>%
  row_spec(1:5, background = "#ffffff", font_size=20) %>%
  column_spec(1:4, color="#616161")
```
<br>
.small-font[
RB: Roger Beecham, JG: Jiaqi Ge, NM: Nik Malleson, RO: Rachel Oldroyd
]

---

## Assessment


```{r outline-lectures-ass-prog, echo=FALSE, out.width="100%"}
outline <- tibble::tibble(
  session = c("1", "2", "3", "4", "5"),
  wc =c("21 Jun", "28 Jun", "5 Jul", "12 Jul", "19 Jul"),
  academic=c("RB","RB", "JG/NM", "RO/RB", "RB/JG/NM/RO"),
  lecture = c("Predictive analytics & microsimulation", "Response modelling and targeted marketing", "Behavioural and agent-based models",
  "Guest lecture", "Wrap-up"),
  deadline=c("","","","Ass #1","Ass #2"),
  progress=c("#1 data","#1 analysis","#2 material","#1 surgery","#2 surgery")

)
kbl(outline) %>%
  row_spec(0, background = "#ffffff", color="#616161", font_size = 18 ) %>%
  row_spec(1:5, background = "#ffffff", font_size=20) %>%
  column_spec(1:4, color="#616161")
```
<br>
.small-font[
RB: Roger Beecham, JG: Jiaqi Ge, NM: Nik Malleson, RO: Rachel Oldroyd
]

---


## How

<img src = "img/why_r.png", width = 60%></img>


---

## R Resources

.small-font[
* H. Wickham and G. Grolemund, R for Data Science, O’Reilly Media, 2017.
  + *The primer for doing data analysis with R.*
  + [Free online](https://r4ds.had.co.nz).

* R. Lovelace et al., Geocomputation with R, CRC Press, 2019.
  + *Comprehensively introduces spatial data handling in R.*
  + [Free online](https://geocompr.robinlovelace.net).

* K. Healy, Data Visualization: A Practical Introduction, Princeton University Press, 2018.
  + *Integrates ggplot2 code with key Information Visualization theory and using real social science datasets.*
  + [Free online -- draft version](https://socviz.co).
]
---

## Module Resources

* [Module website](https://www.roger-beecham.com/predictive-analytics/)
<br><br>
* [Module handbook](https://minerva.leeds.ac.uk/webapps/blackboard/content/listContentEditable.jsp?content_id=_7983617_1&course_id=_521220_1&mode=reset)
<br><br>
* [Assignment briefs](https://minerva.leeds.ac.uk/webapps/blackboard/content/listContentEditable.jsp?content_id=_7983616_1&course_id=_521220_1&mode=reset)
<br><br>
* [Assignment submission page](https://minerva.leeds.ac.uk/webapps/blackboard/content/listContentEditable.jsp?content_id=_7983621_1&course_id=_521220_1&mode=reset)

---

## Module Website

--

<img src = "img/web1.png", width = 40%, style = "position:absolute; top: 20%; left: 8%; box-shadow: 3px 5px 3px 1px #00000080;"></img>

--

<img src = "img/web2.png",  width = 45%, style = "position:absolute; top: 30%; left: 18%; box-shadow: 3px 5px 3px 1px #00000080;"></img>


--

<img src = "img/web3.png",  width = 38%, style = "position:absolute; top: 40%; left: 28%; box-shadow: 3px 5px 3px 1px #00000080;"></img>

---

## How to Learn
--

* *Come to "remote" lectures and labs* <br>
--

* *Engage: Try stuff out, ask questions* <br>
--

* *Contribute to the module [Slack](https://predictive-analytics.slack.com)* <br>
--

* *Coursework throughout: lectures and labs*

---

class: center middle

# Introduction to Predictive Analytics

---


## Data-driven science

**"big data"** and **"data science"** on Google Trends, Oct 2019

<img src = "img/bigdata_trends.png", width = 70%, style = "position:relative; top: 25%; left: 0%; box-shadow: 3px 5px 3px 1px #00000080;"></img>


---

## Data-driven science

.small-font[
* 1000 years ago -- **experimental science**
	+ Description of natural phenomena

* 100s years ago -- **theoretical science**
	+ Newton’s laws, Maxwell’s Equations

* <50 years ago -- **computational science**
	+ Simulate complex phenomena

* Today -- **data-intensive science**
  + Generate knowledge through 	observation
]


<img src = "img/fourth_paradigm.jpeg", width = 35%, style = "position:absolute; top: 25%; left: 55%; box-shadow: 3px 5px 3px 1px #00000080;"></img>

---

## Data-driven science

.small-font[
> *The next generation of scientific discovery will be data-driven as previously unrecognised patterns are discovered by analysing massive and mixed datasets.*
>
> David Willets MP, 2013, Then Minister for Universities and Science
]
--

<img src = "img/end_of_theory.png", width = 50%, style = "position:absolute; top: 45%; left:10%;"></img>



---

## Data-driven science : example

<div class="embed-responsive embed-responsive-16by9">
<iframe width="500" height="350" class="embed-responsive-item" src="https://www.youtube.com/embed/6111nS66Dpk" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
</div>

---

## Data-driven science : example

<img src = "img/google_flu_traps.png", width = 60%></img>

[`Lazer et al. 2014`](https://science.sciencemag.org/content/343/6176/1203.full)

---


## Data-driven science : example

<img src = "img/epj_title.png", width = 40%, style = "position:relative; top: 5%;"></img>

<img src = "img/epj_pic.png", width = 40%, style = "position:relative; top: 15%;"></img>


---

## Data-driven science : example

.small-font[
[`fivethirtyeight.com/parole-assessment-simulator/`](https://projects.fivethirtyeight.com/parole-assessment-simulator/)
]

<img src = "img/fivethirtyeight.png", width = 11%, style = "position:absolute; top: 30%; left: 8%;"></img>


<img src = "img/science_of_sentencing.png", width = 37%, style = "position:absolute; top: 30%; left: 18%;"></img>


---


## Data-driven science : this module

<br>
.pull-left[.small-font[
**Data mining and machine learning** — <br>
Detect hidden patterns in data


**Information Visualization** — <br>
Explore complex structure and patterns in data


**Predictive analytics** — <br>
Predict, under uncertainty, what will happen in future
]
]

<img src = "img/geog5927.png", width = 26%, style = "position:absolute; top: 30%; left: 55%;"></img>

---


## Data-driven science : assignments

<br>
.pull-left[.small-font[
.small-font[
**Assignment 1 -- ** <br> Generate a large synthetic dataset of customers and look for behavioural and demographic associations between individuals to better *target* marketing activity.

<br>

**Assignment 2 -- ** <br> Use data and heuristics to explore and predict how customers behave and respond to different store formats.
]]
]

<img src = "img/geog5927.png", width = 26%, style = "position:absolute; top: 30%; left: 55%;"></img>


---
class: center middle

# Session 1 : Simulating Behaviour

---


## Simulating behaviour practical


<img src = "img/sim_behaviour_prac.png", width = 60%, style = "position:absolute; top: 30%;"></img>

---


## Simulating behaviour practical: survey dataset

<br>

`individuals.csv`
  `15,189 records`

  ```{r survey-data, echo=FALSE, out.width="100%"}
  team <- tibble::tibble(
    var_name = c("age_band", "income_band", "oac_grp", "uk_airport", "overseas_airport", "satisfaction_overall"),
    var_values = c("a24under, ...", "11-15k, ...", "1,2,3,...", "MAN, DSA, ...","TFS, EFL, ...", "1_poor, ..."),
    var_type = c("demographic", "demographic", "geodemographic", "preference", "preference", "preference/attitude")
  )
  kbl(team) %>%
    row_spec(0, background = "#ffffff", font_size = 18) %>%
    row_spec(1:6, background = "#ffffff", font_size = 16)
  ```

---


## Simulating behaviour practical : use case



<img src = "img/beach.jpeg", width = 50%, style = "position:absolute; top: 30%; left: 10%;"></img>




---


## Simulating behaviour practical : use case



<img src = "img/beach_people.jpeg", width = 50%, style = "position:absolute; top: 30%; left: 10%;"></img>


---
## Simulating behaviour practical : use case

.small-font[
`individuals.csv`
  `15,189 records`
  <br>
  `--------`
]
<img src = "img/geogs.png", width = 90%, style = "position:relative;"></img>


---

## Spatial microsimulation


.pull-left[.right[
`Survey data`
.small-font[
individual-level and rich in detail <br>
small sample and may be biased
]]
]

<img src = "img/icon_survey.jpg", width = 20%, style = "position:absolute; top: 28%; left: 52%;"></img>


<br><br><br><br><br><br><br>

.pull-right[.left[
`Census data`
.small-font[
high-level and low in detail <br>
population-level and complete
]]
]

<img src = "img/icon_census.jpeg", width = 20%, style = "position:absolute; top: 65%; left: 28%;"></img>




---

## Spatial microsimulation

<br><br><br>
.small-font[
> *The creation, analysis and modelling of individual-level data allocated to geographic zones.*
>
> Lovelace & Dumont 2016
]


---

## Spatial microsimulation

<br><br><br>
<img src = "img/individuals_areas.jpg", width = 50%, style = "position:relative"></img>

---



## Spatial microsimulation

<br><br><br>
<img src = "img/individuals_areas_constraints.jpg", width = 60%, style = "position:relative"></img>

---

## Spatial microsimulation

<br><br><br><br><br><br>
<img src = "img/individuals_areas_2.jpg", width = 70%, style = "position:relative"></img>

---


## Spatial microsimulation

<br><br><br><br>
<img src = "img/individuals_areas_consraints_2.jpg", width = 75%, style = "position:relative"></img>

---

## Spatial microsimulation

<br>
.pull-left[.right[
Microsimulation does not<br>
generate **new data** <br>
`-----`
]]

--

<br><br>
.pull-right[.left[
`-----` <br>
But **copies of existing data**
]]
---


## Spatial microsimulation: Examples

Health : Smoking <br>
  .small-font[[Tomintz et al. 2008](https://rgs-ibg.onlinelibrary.wiley.com/doi/abs/10.1111/j.1475-4762.2008.00837.x)]

--
.small-font[
 .pull-left[.right[
    **Why?** <br>
    Reported in individual surveys, but not population-level and not from place-to-place
    ]]]
--

  .small-font[.pull-right[.left[
    **Benefits** <br>
    Could be used to target/locate smoking support clinics
  ]]]
--
  .small-font[.pull-right[.left[
    **"Benefits"** <br>
    Could be used by a Tobacco company for targeting investment
  ]]
  ]
---

## Spatial microsimulation: Examples

Economics : Policy Evaluation <br>
  .small-font[[De Agostini et al. 2016](https://www.econstor.eu/bitstream/10419/197592/1/868840475.pdf)]

--
.small-font[
 .pull-left[.right[
    **Why?** <br>
    Simulate / spread impacts inferred from individual-level data over an entire country
    ]]]

--
  .small-font[
  .pull-right[.left[
    **Benefits** <br>
    Quantify (under uncertainty) the impacts of a regressive welfare reform at the country-level

  ]]]

--
  .small-font[.pull-right[.left[
    **"Benefits"** <br>
    Evidence-based decision-making
  ]]]
---


## Spatial microsimulation: Examples

Transport : Simulating travel behaviour <br>
  .small-font[[Lovelace et al. 2014](https://www.sciencedirect.com/science/article/pii/S0966692313001361)]

--
.small-font[
 .pull-left[.right[
    **Why?** <br>
    When designing infrastructure, want to know about the distribution of individuals meeting a particular set of characteristics

    ]]]
--
.small-font[
  .pull-right[.left[
    **Benefits** <br>
    provide evidence around likely winners and losers of a new infrastructure investment

  ]]
  ]
---


## Spatial microsimulation: Assumptions

 .pull-left[
.small-font[
1. Individual-level microdata are representative of the study area <br>

2. Target variable is dependent on the constraint variables in a way that is relatively constant over space and time <br>

3. Input microdataset and constraints are sufficiently rich and detailed to reproduce the full diversity of individuals and areas in the study region
]]

---


## Simulating behaviour practical

 .pull-left[
 <br>
<img src = "img/sim_behaviour_prac.png", width = 100%, style = "position:relative; top: 40%;"></img>
]

 .pull-right[
 <br>
 .small-font[
`individuals.csv : 15,189 records`
  <br>
  `-------` <br>
  spatial microsimulation <br>
  `-------` <br>
  `simulated_oac_age_sex.csv : 320,596 records`
]]
---


## Module Schedule

```{r schedule-again, echo=FALSE, out.width="100%"}
kbl(schedule) %>%
  row_spec(0, background = "#ffffff", font_size = 18 ) %>%
  row_spec(1:5, background = "#ffffff", font_size = 20) %>%
  row_spec(2:5, color="#616161")
```
